---
title: "Reproducible Research Course Project 1"
author: "Sha Li"
date: "December 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###1. Code for reading in the dataset and/or processing the data

```{r}
library(ggplot2)
activity<-read.csv("activity.csv")
```

###2. Histogram of the total number of steps taken each day

Calculate total steps taken per day, missing value is ignored.

```{r}
totalSteps<-aggregate(activity$steps, list(Date=activity$date), FUN=sum)
totalSteps$Date<-as.Date(totalSteps$Date, format="%Y-%m-%d")
```

Plot a histogram of total steps taken per day.

```{r, message=FALSE, warning=FALSE}
qplot(x, data = totalSteps, xlab = "Total Steps Per Day", ylab = "Count", main = "Histogram of Total Steps Per Day")
dev.copy(png, file="figure1.png")
dev.off()
```

###3. Mean and median number of steps taken each day

```{r}
summary(totalSteps)
meanSteps<-mean(totalSteps$x, na.rm = TRUE)
medianSteps<-median(totalSteps$x, na.rm = TRUE)
```

The mean steps per day is `r meanSteps`, the median steps per day is `r medianSteps`.

###4. Time series plot of the average number of steps taken

Plot time series graph that shows the average daily activity pattern, missing value is ignored.

```{r, message=FALSE, warning=FALSE}
meanPattern<-aggregate(activity$steps, list(Interval=activity$interval), FUN=mean, na.rm=TRUE)
ggplot(meanPattern, aes(Interval, x)) + geom_line() + xlab("Interval") + ylab("Averag Steps")+labs(title="Time Series of Average Steps")
dev.copy(png, file="figure2.png")
dev.off()
```

###5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
summary(meanPattern)
intervalMax<-meanPattern[which.max(meanPattern$x),1]
```

Maximum number of steps 206.170 can be found when the 5-minute interval is `r intervalMax`. 

###6. Code to describe and show a strategy for imputing missing data
```{r}
summary(activity$steps)
missing<-sum(is.na(activity$steps))
```

There are `r missing` NAs in the dataset.

Calculate the mean steps based on the 5-minute interval and replace NAs with the mean to create a new dataset called "activityNew".

```{r}
meanDaily<-mean(activity$steps, na.rm = TRUE)
activityNew<-as.data.frame(activity)
activityNew[is.na(activityNew)]<-meanDaily
```

A new data set is created with NAs replaced by `r meanDaily`.

###7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, message=FALSE, warning=FALSE}
totalStepsNew<-aggregate(activityNew$steps, list(Date=activityNew$date), FUN=sum)
totalStepsNew$Date<-as.Date(totalStepsNew$Date,format="%Y-%m-%d")
qplot(x,data = totalStepsNew,xlab = "Total Steps Per Day", ylab = "Count", main = "Histogram of Total Steps Per Day (New)")
dev.copy(png, file="figure3.png")
dev.off()
```

Calculate the new mean and median after imputing NAs.

```{r}
summary(totalStepsNew)
meanStepsNew<-round(mean(totalStepsNew$x))
medianStepsNew<-round(median(totalStepsNew$x))
```

The new mean of total steps taken per day after imputing missing data is `r meanStepsNew`, the new median is `r medianStepsNew`. Both numbers almost remain the same after imputing NAs.

###8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Create a new factor variable in the dataset "activityNew" with two levels: "weekday" and "weekend".

```{r}
dayEnd<-activityNew$date
dayEnd<-weekdays(as.Date(activityNew$date))
dayEnd[dayEnd %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")]<-"Weekday"
dayEnd[dayEnd %in% c("Saturday","Sunday")]<-"Weekend"
dayEnd<-factor(dayEnd, labels = c("Weekday","Weekend"))
activityNew<-cbind(activityNew,dayEnd)
```

Calculate the average steps taken across all weekdays or weekends, plot two time series graphs on the condition of weekdays and weekends.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
meanPatternNew<-aggregate(activityNew$steps, list(Weekday=activityNew$dayEnd, Interval=activityNew$interval), FUN=mean)
ggplot(meanPatternNew, aes(Interval, x)) + geom_line() + xlab("Interval") + ylab("Averag Steps")+labs(title="Time Series of Average Steps")+facet_grid(Weekday~.)
dev.copy(png, file="figure4.png")
dev.off()
```
The graphs show the difference between weekday and weekend steps pattern, weekday has higher max steps than weekend during a specific time interval, however, weekeday has relatively less steps than weekend during other time intervals.
